---
title: "Time Value Money"
output: html_notebook
---


## Inputs


```{r}
library(FinCal)

cape = 26.40
bonds = 0.015
inflation = 0.0145

age.max = 100
age.now = 38
years.life = age.max-age.now
years.kids = 10
```

## Rates

* https://www.morningstar.com/etfs/arcx/tfi/quote
* https://fred.stlouisfed.org/series/T5YIFRM
* https://www.multpl.com/shiller-pe

```{r}
# real composite rate, based on 80/20 allocation
# I'm counting my extra cash as bonds, because they have the same yield right now
alloc_long = 0.9
rate_long = 1/cape*alloc_long + (bond-inflation)*(1-alloc_long)

# betterment recommends 50/50 for a 5y purchase (not sure if that's the same as income over 10y)
alloc_short = 0.5
rate_short = 1/cape*(alloc_short) + (bond-inflation)*(1-alloc_short)

rates = data.frame()
rates.inflation = 0.0145
rates.real = rate_long
rates.real_short = rate_short
rates.bonds = bonds

rates["inflation", "rate"] = rates.inflation
rates["bonds", "rate"] =  rates.bonds
rates["stocks", "rate"] =  1/cape
rates["real long", "rate"] =  rates.real
rates["real short", "rate"] =  rates.real_short
rates
```



## Present Value - Assets

Social Security - My estimated benefit

* https://docs.google.com/spreadsheets/d/1X9DyVoeap0Puf6Jg5Hrd4GA-W5a96Xtceb-LJxP5WPM/edit#gid=5672503

```{r}

ss = data.frame(age = c(66,70), amount = c(20865, 25872), row.names = c("Full","Max"))
# I'm not sure how to calculate this, I lifted it from the bogleheads TVM spreadsheet
ss.pv = 171901
ss
```

Net Cash
```{r}
cash = data.frame()
cash["current", "amount"] = 20000+46000
cash["income","amount"] = 170000
cash["sell","amount"] = 17000
cash["taxes","amount"] = -44000
cash["2020", "amount"] = -15000
cash
```

Investments

```{r}
betterment = 916369

# currently invested 90/10 in betterment
invest = data.frame()
invest["betterment", "stocks"] = round(0.9*betterment)
invest["betterment", "bonds"] = round(0.1*betterment)
invest["hsa", "stocks"] = 7000
invest["hsa", "bonds"] = 0

# private shares - should keep up with inflation. I'll take it's current value
invest["private", "stocks"] = 50000
invest["private", "bonds"] = 0
invest


```



## Present Value - Liabilities

Extra Housing: 10 years estimate of how much extra housing I might need. Ages 13, 11, and 7. Only one child left after 7 years. 

```{r}
housing_extra = 12000 * c(1,1,1,1,1,1,1,1,1,1)
housing_extra.pv = pv.uneven(rates.real_short, housing_extra)
housing_extra.pv
```


Child support

```{r}
cs_payments = function(amount, max, years) {
  c(rep(amount, times=years), rep(0, times=max-years))
}

per_kid = (800/3)*12
kid13 = cs_payments(per_kid, 10, 5)
kid11 = cs_payments(per_kid, 10, 7)
kid07 = cs_payments(per_kid, 10, 10)
child_support = kid13 + kid11 + kid07
# child_support = c(rep(9600, times=5), rep(0, times=5))
child_support
child_support.pv = pv.uneven(rates.real_short, child_support)
child_support.pv
```

Budgets

```{r}
budget = data.frame()
budget["living", "amount"] = 24000
budget["health","amount"] = 2000
budget["spend","amount"] = 10000
budget["home","amount"] = 5000
budget.pv = pv.annuity(rates.real, years.life, sum(budget$amount))
budget
```

```{r}
liabilities = data.frame()
liabilities["child support", "pv"] = child_support.pv
liabilities["housing extra", "pv"] = housing_extra.pv
liabilities["income", "pv"] = budget.pv
liabilities
```


# Analysis, withdrawals, and funded ratio

Withdrawal Rates. Separate pools for kid liabilities because they last 10y.

```{r}
rates.income = pmt(rates.real, years.life, -1, 0, 1)
rates.kids = pmt(rates.real, years.kids, -1, 0, 1)
data.frame(
  rate = c(rates.income, rates.kids),
  row.names = c("Income", "Kids")
)
```


Withdrawal Amounts for Income and Kids.

* [Funded Ratio](https://www.bogleheads.org/forum/viewtopic.php?t=219878)

```{r}

# all assets vs liabilities
total.assets = sum(invest$stocks) + sum(invest$bonds) + sum(cash$amount) + ss.pv
total.liabilities = sum(liabilities$pv)

# if you set aside kid liabilities and deplete them within 10y
kids.liabilities = child_support.pv + housing_extra.pv
kids.assets = -kids.liabilities
kids.withdrawal = rates.kids * kids.assets
kids.funded = kids.assets / -kids.liabilities

# remaining assets support income for life
income.assets = total.assets + kids.liabilities
income.liabilities = budget.pv
income.withdrawal = rates.income * income.assets
income.funded = income.assets / -income.liabilities

data.frame(
  "assets" = c(kids.assets, income.assets),
  "liabilities" = c(kids.liabilities, income.liabilities),
  "withdrawal" = c(kids.withdrawal, income.withdrawal),
  "funded ratio" = c(kids.funded, income.funded),
  row.names = c("Next 10y Kid Expenses ", "Lifetime Income")
)
```



## Allocation

```{r}
# this shows 75/25 right now, including cash as bonds, I need to increase to 80/20 to get enough growth

stock_amount = sum(invest$stocks)
bond_amount = sum(invest$bonds) + sum(cash$amount)
total_amount = stock_amount + bond_amount

short.total = kids.assets
short.stocks =short.total * alloc_short
short.bonds = short.total * (1-alloc_short)

# includes ss, so remove from the allocation calculation
long.total = income.assets - ss.pv
long.stocks = long.total * alloc_long
long.bonds = long.total * (1-alloc_long)

combined.stocks = short.stocks + long.stocks
combined.bonds = short.bonds + long.bonds
combined.total = combined.stocks + combined.bonds


# +/- stocks should equal bonds, and total should be 0
data.frame(
  stocks = c(short.stocks, long.stocks, combined.stocks, stock_amount, stock_amount - combined.stocks),
  other = c(short.bonds, long.bonds, combined.bonds, bond_amount, bond_amount - combined.bonds),
  total = c(short.total, long.total, combined.total, total_amount, total_amount - combined.total),
  allocation = c(short.stocks / short.total, long.stocks / long.total, combined.stocks / combined.total, stock_amount / total_amount, NA),
  row.names = c("Short", "Long", "Combined", "Current", "+/-")
)

# oh, wait, but I don't HAVE the SS yet. That's weird
# it comes out of income, and matches whatever allocation I use
# 120k more into STOCKs, not just betterment. Get spending.
# how about 10k a month? 5k per every 15 days. Sure.


```
